---
title: "Color palette tests for SeqPlots"
author: "Przemyslaw Stempor"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  ioslides_presentation:
    fig_retina: null
    smaller: yes
    transition: 0
widescreen: yes
---


<style>
    .title-slide {
        background-color: #CBE7A5; /* #EDE0CF; ; #CA9F9D*/
    }
    
    slide {
        background-color: #FFFFFF; /* #EDE0CF; ; #CA9F9D*/
    }
    article.smaller pre {
        font-size: 12px;
        line-height: 13px;
        letter-spacing: 0;
    }
    
    slides > slide > hgroup + article {
        margin-top: 10px;
    }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
options("bitmapType" = "cairo")
require(RColorBrewer)
ttl <- 'plot'
```

## ColorBrewer palettes
```{r all_pal, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
display.brewer.all()
```

## ColorBrewer palettes div
```{r div_pal, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
display.brewer.all(type = 'div')
```

## ColorBrewer palettes qual
```{r qual_pal, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
display.brewer.all(type = 'qual')
```

## ColorBrewer palettes seq
```{r seq_pal, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
display.brewer.all(type = 'seq')
```

```{r data, include=FALSE}
require(seqplots)
require(magrittr)
set.seed(1)

source('fn_plotTopBar.R'); plotTopBar <- plotTopBar
devtools::load_all("/Users/przemol/code/seqplotsR")

tracks <- dir('tracks', pattern = 'bw', full.names = TRUE)
features <- dir('features', pattern = 'bed', full.names = TRUE)

ms <- MotifSetup()
ms$addBigWig(tracks[1])$addBigWig(tracks[2])$addBigWig(tracks[3])$addBigWig(tracks[4])
ms$addMotif('CG', revcomp = FALSE, name = "CpG")

Ma <- getPlotSetArray(
    tracks = ms,
    features = features[1],
    refgenome = 'hg19',
    bin = 10L, xmin = 1000L, xmax = 1500L
)

for(i in 1:4) Ma$data[[1]][[i]]$heatmap[is.na(Ma$data[[1]][[i]]$heatmap)] <- 0
Ma$data[[1]][[5]]$heatmap  %>% rowSums  %>% order(decreasing = TRUE) -> ord
for(i in 1:5) Ma$data[[1]][[i]]$heatmap <- Ma$data[[1]][[i]]$heatmap[ord, ]
M <- Ma[1,c(2,1,4,5)]
```

## Current default palette
```{r heatmap1, message=FALSE, paper = 'a4r', warning=FALSE, echo=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
set.seed(1); cls <- plotHeatmap(
    M, labels = colnames(M$as.array())  %>% strsplit('_')  %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'no', clusters = 2,
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22
)
```

# R palettes

## "topo.colors" palette
```{r message=FALSE, paper = 'a4r', warning=FALSE, echo=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
set.seed(1); cls <- plotHeatmap(
    M, labels = colnames(M$as.array())  %>% strsplit('_')  %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'no', clusters = 2,
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
    clspace = topo.colors(11)
)
```

## "terrain.colors" palette
```{r message=FALSE, paper = 'a4r', warning=FALSE, echo=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
set.seed(1); cls <- plotHeatmap(
    M, labels = colnames(M$as.array())  %>% strsplit('_')  %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'no', clusters = 2,
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
    clspace = terrain.colors(11)
)
```

## "heat.colors" palette
```{r message=FALSE, paper = 'a4r', warning=FALSE, echo=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
set.seed(1); cls <- plotHeatmap(
    M, labels = colnames(M$as.array())  %>% strsplit('_')  %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'no', clusters = 2,
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
    clspace = heat.colors(11)
)
```

# ColorBrewer palettes


```{r echo=FALSE, results='asis', message=FALSE, message=FALSE, paper = 'a4r', warning=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
# knit the first three lines of first.Rmd
pal <- RColorBrewer::brewer.pal.info
for (i in 1:nrow(pal)) {
    message(i, 'fdw')
    cat('## Palette', rownames(pal[i,]), ', type=', as.character(pal[[i,2]]), '\n' )
    set.seed(1); cls <- plotHeatmap(
    M, labels = colnames(M$as.array())  %>% strsplit('_')  %>% sapply('[[', 1),
        raster = TRUE, include = c(T,F,F,F), sortrows = 'no', clusters = 2,
        cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
        clspace = RColorBrewer::brewer.pal(pal[i,1], rownames(pal[i,]))
        #o_min = c(-1, -2, -1, 8), #o_max = c(3, 8, 4, 16)
    )
    cat('\n\n')
}
```

```{r echo=FALSE, results='asis', message=FALSE, message=FALSE, paper = 'a4r', warning=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
# knit the first three lines of first.Rmd
pal <- RColorBrewer::brewer.pal.info
for (i in 1:nrow(pal)) {
    message(i, 'rev')
    cat('## Palette: reverced', rownames(pal[i,]), ', type=', as.character(pal[[i,2]]), '\n' )
    set.seed(1); cls <- plotHeatmap(
    M, labels = colnames(M$as.array())  %>% strsplit('_')  %>% sapply('[[', 1),
        raster = TRUE, include = c(T,F,F,F), sortrows = 'no', clusters = 2,
        cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
        clspace = rev(RColorBrewer::brewer.pal(pal[i,1], rownames(pal[i,])))
        #o_min = c(-1, -2, -1, 8), #o_max = c(3, 8, 4, 16)
    )
    cat('\n\n')
}
```
