---
title: "Proposed SeqPlots paper figures - *H. sapiens*"
author: "Przemyslaw Stempor"
date: "July 30, 2015"
output:
  ioslides_presentation:
    fig_retina: null
    smaller: yes
    transition: 0
widescreen: yes
---


<style>
    .title-slide {
        background-color: #CBE7A5; /* #EDE0CF; ; #CA9F9D*/
    }
    
    slide {
        background-color: #FFFFFF; /* #EDE0CF; ; #CA9F9D*/
    }
    article.smaller pre {
        font-size: 12px;
        line-height: 13px;
        letter-spacing: 0;
    }
    
    slides > slide > hgroup + article {
        margin-top: 10px;
    }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
options("bitmapType" = "cairo")
```

## Data aquisition

- Tracks comes from ENCODE project and were downloaded from GEO
- For stronger signal replicates are combined using "bigWigMerge" tool
- Combined tracks are z-scored using bash script with mean and sd calculated by "bigWigInfo" tool"

Example of the data pre-processing;
```{bash, eval=FALSE}
bigWigMerge GSM803485_hg19_wgEncodeHaibTfbsGm12878Pol24h8Pcr1xRawRep2.bigWig \
GSM803485_hg19_wgEncodeHaibTfbsGm12878Pol24h8Pcr1xRawRep1.bigWig \
GSM803355_hg19_wgEncodeHaibTfbsGm12878Pol2Pcr2xRawRep2.bigWig \
GSM803355_hg19_wgEncodeHaibTfbsGm12878Pol2Pcr2xRawRep1.bigWig merge.bedGraph

bedGraphToBigWig merge.bedGraph hg19.chrom.sizes merge.bw

mi=$(bigWigInfo merge.bw | grep mean | sed -n 's/mean: //pg')
sd=$(bigWigInfo merge.bw | grep std | sed -n 's/std: //pg')
cat merge.bedGraph | awk -v mi=$mi -v sd=$sd '{$4=($4-mi)/sd;print}' > zsc.bedGraph    
bedGraphToBigWig zsc.bedGraph hg19.chrom.sizes Pol2_bigWigMerge_zsc.bw

rm merge* zsc.bedGraph
```


## Data aquisition

- Tracks comes from ENCODE project and were downloaded from GEO
- For stronger signal replicates are combined using "bigWigMerge" tool
- Combined tracks are z-scored using bash script with mean and sd calculated by "bigWigInfo" tool"

Example of the data pre-processing;
```{bash, eval=FALSE}
bigWigMerge GSM803485_hg19_wgEncodeHaibTfbsGm12878Pol24h8Pcr1xRawRep2.bigWig \
GSM803485_hg19_wgEncodeHaibTfbsGm12878Pol24h8Pcr1xRawRep1.bigWig \
GSM803355_hg19_wgEncodeHaibTfbsGm12878Pol2Pcr2xRawRep2.bigWig \
GSM803355_hg19_wgEncodeHaibTfbsGm12878Pol2Pcr2xRawRep1.bigWig merge.bedGraph

bedGraphToBigWig merge.bedGraph hg19.chrom.sizes merge.bw

mi=$(bigWigInfo merge.bw | grep mean | sed -n 's/mean: //pg')
sd=$(bigWigInfo merge.bw | grep std | sed -n 's/std: //pg')
cat merge.bedGraph | awk -v mi=$mi -v sd=$sd '{$4=($4-mi)/sd;print}' > zsc.bedGraph    
bedGraphToBigWig zsc.bedGraph hg19.chrom.sizes Pol2_bigWigMerge_zsc.bw

rm merge* zsc.bedGraph
```

## Loading data and libraries
```{r aquire1, message=FALSE}
require(seqplots)
require(magrittr)
set.seed(1)

source('fn_plotTopBar.R'); plotTopBar <- plotTopBar
devtools::load_all("/Users/przemol/code/seqplotsR")

tracks <- dir('tracks', pattern = 'bw', full.names = TRUE)
features <- dir('features', pattern = 'bed', full.names = TRUE)

top_limits <- list(c(-0.25, .1), c(-0.3, 5), c(0, 14), c(0, 25))
top_limits <- NULL
colvec <- list(
    NA, 
    c('black', 'red', 'orange'), 
    c('white', 'grey', 'black'),
    c('white', '#66ccff', '#000066')
)

colvec2 <- list(
    NA, 
    c('black', 'red', 'orange'), 
    c('darkgreen', 'yellow3', 'red'),
    c('white', '#66ccff', '#000066')
)

tracks
features
```

## Data process
```{r aquire2, message=FALSE}
m1 <- getPlotSetArray(
    tracks = grep('H2A.Z|H3k4me3|H3K9ac', tracks, value = TRUE),
    features = grep('band1', features, value = TRUE),
    refgenome = 'hg19',
    bin = 10L, xmin = 1500L, xmax = 3000L
)

m2 <- getPlotSetArray(
    tracks = grep('H3K4me3', tracks, value = TRUE, ignore.case = TRUE),
    features = grep('1|3|5', features, value = TRUE),
    refgenome = 'hg19',
    bin = 10L, xmin = 1500L, xmax = 3000L
)
#MotifSetup()$addMotif('CG', revcomp = FALSE) %>% 
#    getPlotSetArray(features[1], refgenome = 'hg19', bin = 10L, xmin = 1500L, xmax = 3000L) -> clc

ms <- MotifSetup()
ms$addBigWig(tracks[1])$addBigWig(tracks[2])$addBigWig(tracks[3])$addBigWig(tracks[4])
ms$addMotif('CG', revcomp = FALSE, name = "CpG")

Ma <- getPlotSetArray(
    tracks = ms,
    features = features[1],
    refgenome = 'hg19',
    bin = 10L, xmin = 1000L, xmax = 1500L
)

for(i in 1:4) Ma$data[[1]][[i]]$heatmap[is.na(Ma$data[[1]][[i]]$heatmap)] <- 0
Ma$data[[1]][[5]]$heatmap  %>% rowSums  %>% order(decreasing = TRUE) -> ord
for(i in 1:5) Ma$data[[1]][[i]]$heatmap <- Ma$data[[1]][[i]]$heatmap[ord, ]


```

#Figures

## [Plot A](fa2.pdf)

```{r lineplot1, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=8.27, fig.width=11.7, out.height=450, out.width="auto"}
plotAverage(
    m1, labels = colnames(m1$as.array())  %>% strsplit('_')  %>% sapply('[[', 1), 
    legend_pos = 'topleft', legend_ext = TRUE, legend_ext_pos = 'topright',
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
    main = "Top expression quintile", xlab = "TSS", ylab = "Z-scored ChIP signal",
    plotScale = "", #'zscore', 'log2'
)

dev.print(pdf, file='fa2.pdf'); dev.print(png, file='fa2.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Plot B](fb2.pdf)

```{r lineplot2, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=8.27, fig.width=11.7, out.height=450, out.width="auto"}
plotAverage(
    m2, labels = c("TSS_top_20%", "TSS_middle_20%", "TSS_bottom_20%"), 
    legend_pos = 'topleft', legend_ext = TRUE, legend_ext_pos = 'topright',
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
    main = "Top expression quintile", xlab = "TSS", ylab = "Z-scored ChIP signal"
)
dev.print(pdf, file='fb2.pdf'); dev.print(png, file='fb2.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Plot C](fc.pdf)
```{r topbar1, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=2*4, fig.width=6*4, out.height='300', out.width="auto", cache.rebuild=TRUE}
source( 'fn_plotTopBar.R' ); plotTopBar <- plotTopBar
M <- Ma[1,c(2,1,4,5)]
plotTopBar(M, M$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1), top_limits)
dev.print(pdf, file='fc.pdf'); dev.print(png, file='fc.png', height=2*4, width=6*4, res=72, units = "in");
```

## [Plot D](fd.pdf)
```{r heatmap1, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto", cache.rebuild=TRUE}
set.seed(1); cls <- plotHeatmap(
    M, labels = colnames(M$as.array())  %>% strsplit('_')  %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'no', clusters = 2,
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec
    #o_min = c(-1, -2, -1, 8), #o_max = c(3, 8, 4, 16)
)
dev.print(pdf, file='fd.pdf'); dev.print(png, file='fd.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Plot E](fe.pdf)
```{r topbar2, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=2*4, fig.width=6*4, out.height='300', out.width="auto", cache.rebuild=TRUE}
export.bed(import.bed(features[1])[ord][cls$ClusterID==1], 'flt.bed')
MFa <- getPlotSetArray(
    tracks = ms,
    features = 'flt.bed',
    refgenome = 'hg19',
    bin = 10L, xmin = 1000L, xmax = 1500L
)
for(i in 1:4) MFa$data[[1]][[i]]$heatmap[is.na(MFa$data[[1]][[i]]$heatmap)] <- 0
MF <- MFa[1,c(2,1,3,5)]
plotTopBar(MF, MF$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1), top_limits)
dev.print(pdf, file='fe.pdf'); dev.print(png, file='fe.png', height=2*4, width=6*4, res=72, units = "in");
```

## [Plot F](ff.pdf)
```{r heatmap2, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto", cache.rebuild=TRUEpub}
set.seed(1); plotHeatmap(
     MF, labels = MF$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
     raster = TRUE, include = c(F,T,T,T), sortrows = 'decreasing',
     clstmethod = 'ssom', ssomt1 = 1, ssomt2 = 3,
     cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec2,
     o_min = c(0, 0, 0, 0), o_max = c(20, 20, 20, 30)
)

dev.print(pdf, file='ff.pdf'); dev.print(png, file='ff.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Assembled figure (click for full page)](FigHg19L.html#/all) + [Printable - static HTML](static.html)
<iframe src="FigHg19L.html#/all"></iframe>